Open-source landscape for 3D CSEM modelling

### 2020-06-01 manuscript/Manuscript.tex

\mycom[RR]{GMD looks very interesting, my only concern is that it might not be
in the scope of EM modelers, but we can probably reach modeleres from other
communities as well.}

@@

@@

The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of developed numerical
solutions and according publications. The most common applied methods to solve
Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan,
GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and
different variations of the differential equation (DE) method, such as finite
differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin,
GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GJI.11.Schwarzbach,
GEO.04.Commer, GEO.12.daSilva, GJI.13.Puzyrev, GJI.13.Grayver, SEG.16.Zhang},
and finite volumes (FV) \citep{EM.90.Madsen, ECP.07.Haber, GEO.14.Jahandari,
PIER.01.Clemens, GP.06.Mulder}. There are also many different types of
discretisation, where the most common ones are regular grids (Cartesian,
rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid
\citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids
\citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or
OcTree meshes\citep{ECP.07.Haber}. Very well written overviews about the
different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and
\cite{SG.10.Borner}.

In the last 15 years, the publications with regards to 3D EM modelling grew
tremendously, driven by the ever increasing computing power. One reason why
there are so many reports about this topic results from the variety of solution
techniques for the systems of linear equations (SLE). They can be distinguished
between direct solvers \citep{GEO.09.Streich, GEO.15.Grayver, GP.14.Chung,
GEO.14.Jaysaval, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder,
GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers
\citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid
method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

Probably the main reason for the numerous amount of code developments was
already outlined by \cite{SG.05.Avdeev}, who finished his review with the
following statement: «\emph{The most important challenge that faces the EM
community today is to convince software developers to put their 3-D EM forward
and inverse solutions into the public domain, at least after some time. This
would have a strong impact on the whole subject and the developers would
benefit from feedback regarding the real needs of the end-users.}» Some major
exploration service companies as well as certain research consortia and
research groups have their own 3D CSEM codes, but how about codes that are
open-source? Academic codes were traditionally often available upon request by
authors. Codes distributed that way were often provided with the request to not
share the code, and they often come without a license attached which means that
they are, legally, not open-source. Also, at times there can be significant
hurdles to install a code.

The meaning of the term open-source evolved rapidly in the last two decades.
Today, open-source generally not only means that a code is available (with a
proper license), but also that the code is hosted online, is under version
control, has the possibility to file issues and make pull requests to fix bugs.
These are important requirements for reproducible research «\emph{to provide
the means by which the reader can verify the validity of the results and make
use of them in further research}» \citep{GEO.17.Broggini}. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation, often including examples and tutorials. As
such, the term has shifted from purely open-source codes to developments in the
open with increasingly building a global community. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}.
\imycom[DW]{Others coming to your mind?\imycom[RR]{no thoughts about geophysics,
  but there are many projects for related topics such as gmt, gis, mesh-gen}}

We continue addressing the topic of open-source developments with respect to 3D
CSEM modeling. Even though it comprises still the minority of developments, the
landscape in this field of research changed in the last five years. Before,
there were only closed-source codes owned by companies or  consortia and
individual projects, often without a proper documentation, that could be
requested from the authors. We introduce and compare four recent open-source
projects with focus on 3D CSEM modeling. All presented codes are in the Python
ecosystem and use either the FD method on structured grids or the FE method on
unstructured tetrahedral meshes.

Furthermore, we deal with the topic of verification. \cite{GJI.13.Miensopust}
presents a review of two workshops dealing with the verification of
magnetotelluric forward and inversion codes, but we are not aware of any
comparable study or benchmark suite for CSEM data. Hence, we first conduct a
similar exercise for CSEM codes. We want our efforts to help to ensure the
reliability of simulations, not only carried out with the available open-source
codes, but also for individual closed-source project. Analytical and
semi-analytical solutions only exist for simple halfspace or layered-Earth
models, which served mostly to verify new codes. Beyond them, the only
objective possibility to ensure the accuracy of solutions is by comparing
results from different modellers. If different discretisations and
implementations of Maxwell's equations yield the same result, it gives
confidence in their accuracy.

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}. All four
codes have their user-facing routines written in Python; all of them make heavy
use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy}
\citep{NM.20.Virtanen}. The four of them are “modern” open-source projects,
meaning that they come with both, an open-source license and an online-hosted
version-control system with tracking possibilities (raising issues, filing pull
requests). All developments comprise an extensive online documentation with
many examples and have continuous integration to some degree. Newer
package-management systems such as Conda, Docker, or pip significantly
supported us to make our codes easily available and installable without any
user-side compilations. Each one can be downloaded and installed with a single
command.

In the following a quick rundown of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?} diffusive
approximation, hence neglecting displacement currents. The machines on which
the different codes were run are listed in Table~\ref{tbl:machines}, together
with the responsible operator.

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration.
  Note that \emg3d is a solver on its own, while the other codes implement
  third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar},
  \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

@@

The thin isotropic layer beneath the water requires many tetrahedra for a
good-quality discretisation. Therefore, the FE codes custEM and PETGEM used the
same tetrahedral mesh with a horizontal layered-earth extent of 30\,km from the
origin, surrounded by a ten times larger halfspace-like boundary mesh with
water conductivities assigned to the subsurface. This approximation
significantly reduced the problem size compared to considering the
layered-earth geometry for the whole computational domain. \imycom[RR]{Add
short information about mesh design with SIMPEG and emg3d, including something
like "comparatively coarse discretization in center to minimize computation
times" }

@@

\imycom[RR]{Agreed. Another suggestion, I solved this visualization problem
  with a color coded matrix plot, showing 90th percentiles or median values of
  all errors along the line as meaningful comprehensive number. Look in my
  thesis - example 4. Comparison for real and imag could be in the lower and
  upper triangles, or we show absolute magnitudes for layered earth in lower
  triangle and for block model in upper triangle. With help of this plot, one
  could easily say if differences are higher for block model and which codes
  fit to each other well.}
\imycom[RR]{If using the mentioned matrix plots, we would need only one of the
  figures 4-6, and could show the other data comprehensively. I added a python
  script to generate an example of my plotting suggestion in the model-block
  directory and a new Figure 7 in the manuscript that illustrates what I mean.
  Of course, any upgrade suggestion for this initial idea is welcome, aside
  from making it looking nice with labels  and annotations.}
}

@@

%
\tabl[btp]{comp-block}{Comparison of used CPU and RAM and of number of cells
  used to discretize the block model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 & 24 & 0.0 & 0 \\
  \emg3d  & 0.0 &  1 & 0.0 & 0 \\
  \petgem & 0.0 & 48 & 0.0 & 0 \\
  \simpeg & 0.0 &    & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in
the north-eastern part of the Campos Basin, offshore Brazil, which was
discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well
log data a realistic resistivity model with vertical transverse isotropy (VTI),
Marlim R3D (MR3D), which they released under an open-source license.
\cite{GEO.19.Correa} computed CSEM data for six frequencies from 0.125\,Hz to
1.25\,Hz, released under an open-source license. To compute the data they used
a state-of-the-art code from the industry \citep{GEO.07.Maao}. It is therefore
an ideal case to verify our open-source codes against, as it is a complex,
realistic model and the data were computed by an industry-proofed
code. Additionally, that code is a time-domain code, whereas the four codes
under consideration here compute the results all in the frequency-domain.
\imycom[RR]{I would delete this sentence, not that important and a bit
redundant.\imycom[DW]{Did you refer to the last sentence or the one before? I
think the last sentence contains quite important info.}}

The detailed reservoir model consists of 1022 by 371 by 1229 cells, totalling
to almost 466 million cells, where each cell has dimensions of 25 by 75 by
5\,m. For the computation the model is upscaled to 515 by 563 by 310 cells,
totalling to almost 90 million cells, where each cell has dimensions of 100 by
100 by 20\,m.

@@

We computed all all six frequencies for all components ($E_x$, $E_y$, $E_z$)
for both inline and broadside acquisition lines. However, we show here only the
three frequencies and the responses with the strongest amplitudes (which are
generally used in processing), which are $E_x$ inline and broadside and $E_y$
broadside. The comparison is shown in Figure~\ref{fig:results-marlim}.
%
\plot*{results-marlim}{width=\textwidth}{
  This is one heck of a figure. Just an idea, please make suggestions how to
  improve it (currently, \petgem data is a copy of \custem data, and \simpeg
  data is a copy of \emg3d data).
}
%

@@

%
\tabl[btp]{comp-marlim}{Comparison of used CPU and RAM and of number of cells
  used to discretize the Marlim R3D model. \imycom[DW]{I will fill this table
  out once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %{ls[table-format=4.2]rs[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

We point out that making a synthetic and a real benchmark model available to
cross-validate any 3D CSEM code was the determining reason for their choice,
regardless of the suitability of the considered FD and FE codes for these
problems. The first model, including the block-anomalies variation, is exactly
reproducible by using any numerical method with either regular or irregular
computational domains. Aside from the idea of analyzing an industrially
relevant benchmark example, the Marlim R3D model was also considered since the
primary design from \cite{GEO.19.Correa} uses a regular discretisation. At
least theoretically, the exact geometry might be reconstructed with
unstructured meshes, but it is never possible to convert an irregular geometry
to a regular grid without approximations. Hence, the chosen examples favoured
the finite-difference codes, since they used either regular grids or
octree-meshes, whereas the two finite-element codes were required to
tetrahedralise the model geometries.

@@

\imycom[DW]{I would like to capture as many as possible of the (I) differences,
  (II) difficulties, (III) advantages, and (IV) disadvantages, that we
  encountered in this exercise. This includes differences from direct to
  iterative solver and differences from FE to FD codes etc. Please just jot
  down things that come to your mind.}

@@

The work of D.W. was conducted within the Gitaro.JIM project funded through
MarTERA, a EU Horizon 2020 research and innovation programme (No 728053);
\href{https://www.martera.eu}{martera.eu}. The development of custEM as part of
the DESMEX project was funded by the Germany Ministry for Education and
Research (BMBF) in the framework of the research and development program
Fona-r4 under grant 033R130D.

>>>>

\mycom[RR]{GMD looks very interesting, my only concern is that it might not be
in the scope of EM modelers, but we can probably reach modeleres from other
communities as well.}
\mycom[OCR]{I also find GMD very interesting, in fact it was my first option to publish the first petgem version (2 years ago). The answer was: "manuscript outside of journal's scope". In this sense I agree with Raphael. My vote is for SIG}

@@

\imycom[OCR]{In general, I believe that the content is adequate and the ideas are clear and in accordance with the objective that we have initially set. However, I want to highlight the task of being careful when justifying the absence of the SimPEG results for the marlim model (I have read it in slack, although it may not be definitive). I believe that this may or may not be a weak point depending on how we explain it. I know that perhaps it is not serious, I only make this comment so that the speech justifies the facts. }

@@

\mycom[OCR]{I have rearranged some references according to the year of publication (here and in the rest of the paper)}
The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of developed numerical
solutions and according publications. The most common applied methods to solve
Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan,
GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and
different variations of the differential equation (DE) method, such as finite
differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin,
GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer,
GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang},
and finite volumes (FV) \citep{EM.90.Madsen, PIER.01.Clemens, GP.06.Mulder, ECP.07.Haber, GEO.14.Jahandari}. There are also many different types of
discretisation, where the most common ones are regular grids (cartesian,
rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid
\citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids
\citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or
OcTree meshes\citep{ECP.07.Haber}. Very well written overviews about the
different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and
\cite{SG.10.Borner}.

In the last 15 years, the publications with regards to 3D EM modelling grew
tremendously, driven by the ever increasing computing power. One reason why
there are so many reports about this topic results from the variety of solution
techniques for the systems of linear equations (SLE). They can be distinguished
between direct solvers \citep{GEO.09.Streich, GP.14.Chung,
GEO.14.Jaysaval, GEO.15.Grayver, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers
\citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid
method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

Probably the main reason for the numerous amount of code developments was
already outlined by \cite{SG.05.Avdeev}, who finished his review with the
following statement: «\emph{The most important challenge that faces the EM
community today is to convince software developers to put their 3-D EM forward
and inverse solutions into the public domain, at least after some time. This
would have a strong impact on the whole subject and the developers would
benefit from feedback regarding the real needs of the end-users.}» Some major
exploration service companies, as well as certain research consortia and
research groups, have their own 3D CSEM codes, but how about codes that are
open-source? Academic codes were traditionally often available upon request by
authors. Codes distributed that way were often provided with the request to not
share the code, and they often come without a license attached which means that
they are, legally, not open-source. Also, at times there can be significant
hurdles to install a code.

The meaning of the term open-source evolved rapidly in the last two decades.
Today, open-source generally not only means that a code is available (with a
proper license), but also that the code is hosted online, is under version
control, has the possibility to file issues and make pull requests to fix bugs.
These are important requirements for reproducible research «\emph{to provide
the means by which the reader can verify the validity of the results and make
use of them in further research}» \citep{GEO.17.Broggini}. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation, often including examples and tutorials. As
such, the term has shifted from purely open-source codes to developments in the
open with increasingly building a global community. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}.
\imycom[DW]{Others coming to your mind?\imycom[RR]{no thoughts about geophysics,
  but there are many projects for related topics such as gmt, gis, mesh-gen}
  \imycom[OCR]{No idea from my side}}

We continue addressing the topic of open-source developments with respect to 3D
CSEM modeling. Even though it comprises still the minority of developments, the
landscape in this field of research changed in the last five years. Before,
there were only closed-source codes owned by companies or  consortia and
individual projects, often without proper documentation, that could be
requested from the authors. We introduce and compare four recent open-source
projects with focus on 3D CSEM modeling. All presented codes are in the Python
ecosystem and use either the FD method on structured grids or the FE method on
unstructured tetrahedral meshes.

Furthermore, we deal with the topic of verification. \cite{GJI.13.Miensopust}
presents a review of two workshops dealing with the verification of
magnetotelluric forward and inversion codes, but we are not aware of any
comparable study or benchmark suite for CSEM data. Hence, we first conduct a
similar exercise for CSEM codes. We want our efforts to help to ensure the
reliability of simulations, not only carried out with the available open-source
codes, but also for individual closed-source projects. Analytical and
semi-analytical solutions only exist for simple halfspace or layered-Earth
models, which served mostly to verify new codes. Beyond them, the only
objective possibility to ensure the accuracy of solutions is by comparing
results from different modellers. If different discretisations and
implementations of Maxwell's equations yield the same result, it gives
confidence in their accuracy.

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}. All four
codes have their user-facing routines written in Python; all of them make heavy
use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy}
\citep{NM.20.Virtanen}. The four of them are “modern” open-source projects,
meaning that they come with both, an open-source license and an online-hosted
version-control system with tracking possibilities (raising issues, filing pull
requests). All developments comprise an extensive online documentation with
many examples and have continuous integration to some degree. Newer
package-management systems such as Conda, Docker, or pip significantly
supported us to make our codes easily available and installable without any
user-side compilations. Each one can be downloaded and installed with a single
command.

In the following a quick rundown of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?}\mycom[OCR]{petgem consider the weak work of maxwell equations in their diffusive form (quasi-static approximation)} diffusive
approximation, hence neglecting displacement currents. The machines on which
the different codes were run are listed in Table~\ref{tbl:machines}, together
with the responsible operator.

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration.
  Note that \emg3d is a solver on its own, while the other codes implement
  third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar},
  \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

\petgem is a parallel code for frequency-domain 3D CSEM data from marine and land surveys. The high-order edge FE method (HEFEM) is used for the discretization of the governing equations in its diffusive form. This technique provides a suitable mechanism to obtain stable numerical solutions as well as a good trade-off between number of dof\mycom[OCR]{I assume we have previously mentioned the term dofs} and computational effort. The current implementation supports up to six-order tetrahedral vector basis functions. Moreover, because the HEFEM belongs to the FE family, the unstructured meshes can be used efficiently on complex geometries (e.g., models with topography and bathymetry).

\petgem permits to locate the source and receivers anywhere in the computational domain (e.g., sediments, seafloor, sea, ground), allowing to analyze the physical environment of the electric responses and how parameters impact them (e.g., frequency, conductivity, dependence on mesh setup, basis order, solver type, among others). Furthermore, \petgem implements a semi-adaptive mesh strategy ($hp$ mesh refinement) based on physical parameters and on polynomial order to satisfy quality criteria chosen by the user. Nonetheless, only Horizontal electric dipole (HED) has been implemented.

For the parallel forward modelling computations, a highly scalable MPI domain decomposition allows reducing run-times and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-performance Computing architectures).

Based on the current state of the \petgem project, there are many possibilities
to improve or to add additional features. For instance, in the short-term, more
modeling routines and support for multi-source set-ups will be implemented. For the long-term, the obvious next step is to move to 3D CSEM data inversion.

@@

The thin isotropic layer beneath the water requires many tetrahedra for a
good-quality discretisation. Therefore, the FE codes \custem and \petgem used the
same tetrahedral mesh with a horizontal layered-earth extent of 30\,km from the
origin, surrounded by a ten times larger halfspace-like boundary mesh with
water conductivities assigned to the subsurface. This approximation
significantly reduced the problem size compared to considering the
layered-earth geometry for the whole computational domain. \imycom[RR]{Add
short information about mesh design with SIMPEG and emg3d, including something
like "comparatively coarse discretization in center to minimize computation
times" }

@@

\imycom[RR]{Agreed. Another suggestion, I solved this visualization problem
  with a color coded matrix plot, showing 90th percentiles or median values of
  all errors along the line as meaningful comprehensive number. Look in my
  thesis - example 4. Comparison for real and imag could be in the lower and
  upper triangles, or we show absolute magnitudes for layered earth in lower
  triangle and for block model in upper triangle. With help of this plot, one
  could easily say if differences are higher for block model and which codes
  fit to each other well.}
\imycom[RR]{If using the mentioned matrix plots, we would need only one of the
  figures 4-6, and could show the other data comprehensively. I added a python
  script to generate an example of my plotting suggestion in the model-block
  directory and a new Figure 7 in the manuscript that illustrates what I mean.
  Of course, any upgrade suggestion for this initial idea is welcome, aside
  from making it looking nice with labels  and annotations.}
\imycom[OCR]{I really liked this way of presenting the comparison, in addition to saving space, we gained in understanding. Looks really nice!}

}

@@

%
\tabl[btp]{comp-block}{Comparison of used CPU and RAM and of number of cells
  used to discretize the block model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.\imycom[OCR]{I can help you to fill this table and with its analysis}}}{
  \centering
\begin{tabular}{c c c c c}  %lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 & 24 & 0.0 & 0 \\
  \emg3d  & 0.0 &  1 & 0.0 & 0 \\
  \petgem & 0.0 & 48 & 0.0 & 0 \\
  \simpeg & 0.0 &    & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in
the north-eastern part of the Campos Basin, offshore Brazil, which was
discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well
log data a realistic resistivity model with vertical transverse isotropy (VTI),
Marlim R3D (MR3D), which they released under an open-source license.
\cite{GEO.19.Correa} computed CSEM data for six frequencies from 0.125\,Hz to
1.25\,Hz, released under an open-source license. To compute the data they used
a state-of-the-art code from the industry \citep{GEO.07.Maao}. It is therefore
an ideal case to verify our open-source codes against, as it is a complex,
realistic model and the data were computed by an industry-proofed
code. Additionally, that code is a time-domain code, whereas the four codes
under consideration here compute the results all in the frequency-domain.
\imycom[RR]{I would delete this sentence, not that important and a bit
redundant.\imycom[DW]{Did you refer to the last sentence or the one before? I
think the last sentence contains quite important info. .\imycom[OCR]{If possible, I would like to uphold this sentence. I agree with dieter, it contains relevant information.}}}

The detailed reservoir model consists of 1022 by 371 by 1229 cells\mycom[OCR]{I prefer this notation $1022 \times 371 \times 1229$. Nothing special, just format}, totalling
to almost 466 million cells, where each cell has dimensions of 25 by 75 by
5\,m. For the computation the model is upscaled to 515 by 563 by 310 cells,
totalling to almost 90 million cells, where each cell has dimensions of 100 by
100 by 20\,m.

@@

We computed all all six frequencies for all components ($E_x$, $E_y$, $E_z$)
for both inline and broadside acquisition lines. However, we show here only the
three frequencies and the responses with the strongest amplitudes (which are
generally used in processing), which are $E_x$ inline and broadside, and $E_y$
broadside. The comparison is shown in Figure~\ref{fig:results-marlim}.
%
\plot*{results-marlim}{width=\textwidth}{
  This is one heck of a figure. Just an idea, please make suggestions how to
  improve it (currently, \petgem data is a copy of \custem data, and \simpeg
  data is a copy of \emg3d data).
}
%

@@

%
\tabl[btp]{comp-marlim}{Comparison of used CPU and RAM and of number of cells
  used to discretize the Marlim R3D model. \imycom[DW]{I will fill this table
  out once we all submitted the final versions of the results.\imycom[OCR]{I can help you to fill this table and with its analysis}}}{
  \centering
\begin{tabular}{c c c c c}  %{ls[table-format=4.2]rs[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

We point out that making a synthetic and a real benchmark model available to
cross-validate any 3D CSEM code was the determining reason for their choice,
regardless of the suitability of the considered FD and FE codes for these
problems. The first model, including the block-anomalies variation, is exactly
reproducible by using any numerical method with either regular or irregular
computational domains. Aside from the idea of analyzing an industrially
relevant benchmark example, the Marlim R3D model was also considered since the
primary design from \cite{GEO.19.Correa} uses a regular discretisation. At
least theoretically, the exact geometry might be reconstructed with
unstructured meshes, but it is never possible to convert an irregular geometry
to a regular grid without approximations. Hence, the chosen examples favoured
the FD codes, since they used either regular grids or
octree-meshes, whereas the two FE codes were required to
tetrahedralise the model geometries.

@@

\imycom[DW]{I would like to capture as many as possible of the (I) differences,
  (II) difficulties, (III) advantages, and (IV) disadvantages, that we
  encountered in this exercise. This includes differences from direct to
  iterative solver and differences from FE to FD codes etc. Please just jot
  down things that come to your mind.
  \imycom[OCR]{ I think an important point is the user experience (modeler). I know this is an intangible element but perhaps we could add a line about it. It is my opinion}}

@@

The work of D.W. was conducted within the Gitaro.JIM project funded through
MarTERA, a EU Horizon 2020 research and innovation programme (No 728053);
\href{https://www.martera.eu}{martera.eu}. The development of \custem as part of
the DESMEX project was funded by the Germany Ministry for Education and
Research (BMBF) in the framework of the research and development program
Fona-r4 under grant 033R130D. The work of O.C-R. has received funding from the \textit{European Union\textquotesingle s Horizon 2020 programme} under the \textit{Marie Sklodowska-Curie} grant agreement N$^\circ$ 777778. Further, the development of \petgem has received funding from the \textit{European Union\textquotesingle s Horizon 2020 programme}, grant agreement N$^\circ$ 828947, and from the Mexican Department of Energy, CONACYT-SENER Hidrocarburos grant agreement N$^\circ$ B-S-69926. Furthermore, O.C-R. has been 65\% cofinanced by the European Regional Development Fund (ERDF) through the Interreg V-A Spain-France-Andorra program (POCTEFA2014-2020). POCTEFA aims to reinforce the economic and social integration of the French-Spanish-Andorran border. Its support is focused on developing economic, social and environmental cross-border activities through joint strategies favouring sustainable territorial development.\mycom[OCR]{My apologies for these long acks, but they are the projects in which I am participating and the BSC makes a lot of emphasis on complying verbatim with the thansk sentences (too rigorous in my opinion)}
